---
title: "ANOVA"
output: html_document
date: "2025-05-28"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

We had a look at how ANOVA can be used to compare multiple groups at once. Now
it is time for you to try

 - Replace _ to complete expressions and return the correct answer 
 - Restart R to remove previous results using: Ctrl/Cmd + shift + F10
 - Remember - try to use google to help you answer questions

## Import packages

Import the `tidyverse` for `dplyr` and `ggplot2` packages

```{r}
_
```

## One-way ANOVA

`Turtles.csv` is a study that examined the he hatching times of turtle eggs 
incubated at four different temperatures (15°C, 20°C, 25°C and 30°C). Hatching 
time (Days) is the continuous response variable and temperature is the categorical 
predictor variable with with four levels. 

### Import the data

Import `Turtles.csv` from the `data/` folder

```{r}
turtle_hatching <-
  _
```


### Examine the data 

Count the number of observations per temperature

- Confirm there are 4 temperature levels
- How many replicates are there per temperature? If they are all the same, we called this a 'balanced design'

```{r}
turtle_hatching %>% 
  _
```

### Visualise data

Create a boxplot of days until hatching by temperature groups

 - There should be 4 boxes. 
 - What data type is Temperature currently? We will fix it in the next step
 
```{r} 
turtle_hatching %>% 
  _
```

#### Data transformation 

Turn Temperature into a factor using `mutate()` and `factor()`, then remake the 
boxplot (next section)

```{r}
turtle_hatching <-
  turtle_hatching %>% 
  _
```

#### Recreate the boxplot

```{r} 
turtle_hatching %>% 
  _
```

Export the plot to the outputs folder as a 120 x 120 mm .png

```{r}
ggsave(_)
```

### Conduct an ANOVA

What is the null hypothesis here?

```code
```

Create an ANOVA object using `aov()`

 - Remember to use a formula y ~ x and provide the data argument
 
```{r}
turtle_hatching_aov <-
  _
```

Obtain the traditional ANOVA table using `summary()`

```{r}

```

Is there a difference in hatching time between the 4 temperatures tested?

```text
_
```

### Conduct pairwise comparisons

Use `TukeyHSD()` to conduct pairwise comparisons. Which temperatures differed from each other?

```{r}
_
```

Write a short conclusion paragraph

```{r}

```

```{r, include=F}
# If you have successfully completed the code above, try to "knit" the document"
# Use the "knit" button at the top of the script panel
```

## Extra exercises

### One factor ANOVA - Bag strength

A paper manufacturer makes grocery bags. They are interested in increasing the 
tensile strength of their product. It is thought that strength is a function of
the hardwood concentration in the pulp. An investigation is carried out to compare
four levels of hardwood concentration: 5%, 10%, 15% and 20%. Six test specimens 
are made at each level and all 24 specimens are then tested in random order

```{r, eval = F}
# Import data
bags <-
  readr::read_csv()

# Examine data and prepare
bags %>% 
  count(_)

bags %>% 
  ggplot(aes(x = _, y = _)) +
  _

bags <- 
  bags %>% 
  mutate(_ = factor(_))

bags %>% 
  ggplot(aes(x = _, y = _)) +
  geom_boxplot()

# ANOVA
bags_aov <-
  aov(_ ~ _, data = _)

_ %>% 
  summary()

# Pairwise comparisons 
bags_aov %>% 
  _()
```


### Two factor ANOVA - Battery life

`battery_life.csv` contains data on battery life under different temperature and
materials: The effective life (in hours) of batteries is compared by material type 
(1, 2 or 3) and operating temperature: Low (-10˚C), Medium (20˚C) or High (45˚C).
Twelve batteries are randomly selected from each material type and are then randomly 
allocated to each temperature level.


```{r, eval = F}
# Import battery_life.csv
battery_life <-
  readr::read_csv(_)

# Examine data
battery_life %>% 
  count(_, _)

# Prepare data
# material_type is numeric and needs to be a factor
# temperature requires used to specific the levels of the factor for low, med, high
battery_life <-
  battery_life %>% 
  mutate(material_type = factor(_),
         temperature = factor(_, levels = c(_)))

# Plot data
battery_life %>% 
  ggplot(aes(x=_, y = _, colour = _)) +
  geom_point() +
  stat_summary(mapping = aes(group = _), geom = 'line', fun = 'mean')

# Conduct ANOVA
battery_aov <-
  aov(life_h ~ _ * _, data = _)

_ %>% 
  summary()

# Conduct pairwise comparisons
TukeyHSD(_)
```

