---
title: "Linear regression"
output: html_document
date: "2025-05-28"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

We had a look a simple linear regression. Now lets flex our linear model muscles.

 - Replace _ to complete expressions and return the correct answer
 - Restart R to remove previous results using: Ctrl/Cmd + shift + F10
 - Remember - try to use google to help you answer questions


## Import packages

Import the `tidyverse` for `dplyr` and `ggplot2` packages

```{r}
_
```

### Simple linear regression

A researcher wants to know if flipper length is a good proxy for body mass for 
Adelie penguins. They want to build a model that will allow them to predict body
mass by only measuring flipper length.

### Import the data

`penguins.csv` contains morphometric data on 3 species of penguins including
Adelie penguins. Import the file from the `data/` folder

```{r}
penguins <-
  _
```


### Data preparation

Filter the data to only include the species Adelie

```{r}
penguins_adelie <-
  penguins %>% 
  _
```

### Visualise the data

Create a scatter plot of flipper_length_mm and y = body_mass_g, include a geom_smooth
layer with method = 'lm'

```{r}
penguins_adelie %>% 
  ggplot(aes(x = _, y = _)) +
  _ +
  _
```

In this particular case, with a simple linear regression, `geom_smooth(method = 'lm')`
adds to the plot the results of a simple linear regression. In more complex cases, 
we may need to obtain data from the model and add that as a layer to ggplot. Furthermore,
we do not have the model object to inspect.

Finalise the plot and export it to outputs as a 120 x 120 mm png

```{r}
ggsave(_)
```


### Fit a linear regression model

Use lm() to fit a simple linear regression model.

```{r}
penguins_adelie_lm <-
  lm(_,_ data = _)
```

`penguins_adelie_lm` is a linear model object, and simply a **list** (a data structure
in R that has different elements of data, accessed with `$`, like a column in a 
data.frame) with data elements from the model fit. We generally never have to access the list elements manually as we use accessor functions for that. But for example, I have 
print the structure of the object to show there are names to select various parts
of the model if needed

```{r}
str(penguins_adelie_lm)
```

#### Examine the fitted model

Examine the residuals of the model using a residual vs fits plot (which = 1). Are
they evenly spread around a residual of 0 in a blob type pattern?

```{r}
plot(_, which = 1, pch = 19)
```

There are other diagnostic plots for lm .. we are only looking at 1.

```{r, eval = F}
?plot.lm
```

We can also examine how far each point is from the line of best fit.We should see
a normal, bell shaped curve if the model fits the data well. 

This is one example of us accessing a data element from the `lm()` object. Make
a histogram using `hist()` of the `residuals` element of `penguins_adelie_lm`

```{r}
hist(penguins_adelie_lm$_, breaks = 50)
```


### Obtain a summary of the model ----

Use `summary()` to obtain the model summary

  - Is flipper length a good predictor of body mass?
  - Each unit increase in flipper_length_mm is associated with what change in body_mass_g 

```{r}

```

Write a short conclusion paragraph

```text

```

### Obtaining model predictions

We can use `predict()` to obtain model predictions. We have to provide a data.frame
with values to predict, and the column name must match that of the predictor used
in `lm()`. The data.frame is passed to the `newdata=` argument

Obtain the predicted mass for a flipper_length_mm of 200

```{r}
predict(penguins_adelie_lm, 
        newdata = data.frame(flipper_length_mm = _))
```

Depending on the task we can obtain confidence or prediction intervals around the 
estimates. Confidence intervals are used to estimate parameters, while prediction 
intervals are used to make predictions about future values.

To obtain confidence intervals, we need to include the interval argument
Include 'confidence' in the interval argument below

```{r}
predict(penguins_adelie_lm, 
        newdata = data.frame(flipper_length_mm = 200), 
        interval = _)
```

We can obtain many predictions at once. We need to make a data.frame of values to
use with `predict()`. Create a sequence from 170 to 210 by 10 using `seq()` within
the following code that will create a dta.frame with the column `flipper_length_mm`

```{r}
flipper_values_predict <-
  data.frame(
    flipper_length_mm = _
  )

flipper_values_predict
```

Then pass it to predict, which produced a data.frame with 3 columns - `fit`, `lwr`, `upr`

```{r}
flipper_body_mass_predictions <-
  predict(penguins_adelie_lm, 
          newdata = _, 
          interval = "confidence")

flipper_body_mass_predictions
```

Join back the `flipper_length_mm` values

```{r}
flipper_body_mass_predictions <-
  flipper_body_mass_predictions |> 
  bind_cols(
    flipper_values_predict
  )

flipper_body_mass_predictions
```

You could use the predictions data.frame to add the model data layer to a ggplot()
instead of using geom_smooth(). This is more useful with complex models.

```{r}
penguins_adelie %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_line(data = flipper_body_mass_predictions, mapping = aes(y = fit), linewidth = 1) +
  geom_line(data = flipper_body_mass_predictions, mapping = aes(y = lwr)) +
  geom_line(data = flipper_body_mass_predictions, mapping = aes(y = upr))
```


```{r, include=F}
# If you have successfully completed the code above, try to "knit" the document"
# Use the "knit" button at the top of the script panel
```


## Extra exercises - Multiple linear regression

The cell_growth.csv contains the cell diameter and concentration of Tetrahymena 
cells with and without glucose added to growth medium. Glucose = 1 means glucose
was added, Glucose = 2 means glucose was absent

A researcher wants to know how cell concentration and glucose addition affect
the cell diameter of Tetrahymena

```{r}
# 2.1 Import data ----
# Import cell_growth.csv
cell_growth <-
  readr::read_csv(_)

cell_growth

# 2.2 Visualise data ----
# Make a scatter plot of cell diameter and concentration, coloured by glucose
# The glucose legend should only be 2 values
# Is trend linear?
cell_growth %>% 
  ggplot(aes(x = _ , y = _, color = _)) +
  geom_point()

# We need to create a factor out of glucose
cell_growth <-
  cell_growth %>% 
  mutate(glucose = factor(_, labels = c('Yes', 'No')))

# Try log transforming cell_conc before making the plot
# If the trend look linear, then that is how we should include the predictor in the
# linear model
cell_growth %>% 
  ggplot(aes(x = _(cell_conc) , y = cell_diameter, color = glucose)) +
  geom_point() 


# 2.3 Fit a linear model ----
# We are using the * between predictor for a "crossed" design. It means the
# regression slope for each glucose treatment can be different (we could also 
# constrain them to be the same)
cell_growth_lm <-
  lm(_ ~ log10(_) * _, data = cell_growth) 


# 2.4 Examine the fitted model ----
# Are the residuals evenly spread without nay pattern?
# Possible outliers have numbers, but we will not deal with them further
plot(_, which = 1, pch = 19)

# 2.5 Obtain the summary of the fitted model ----
# Use summary() on the linear model object
#
# This is be a fit difficult to read if you have not seen it before. It is like
# like this is for a statistical estimation reason that we will not discuss further 

# (Intercept)                        = the intercept for glucose = 1
# log10(cell_conc)                   = the slope for glucose = 1             
# factor(glucose)2                   = difference in intercept for glucose 2 to glucose 1   
# log10(cell_conc):factor(glucose)2  = difference in slope for glucose 2 to glucose 1

# Is the slope between cell_diameter and log10(cell_conc) difference between glucose
# levels?
_

# 2.6 Analysis of variance ----
# We can conduct an ANOVA on this model to test the hypotheses
# 1. Is the a relationship between cell_diameter and cell concentration regardless of glucose addition?
# 2. Does glucose addition affect cell diamenter regardless of cell concentration?
# 3. Does glucose affect the relationship between cell diameter and cell concentration?
# Use anova() on the linear model object 
_
```
