---
title: "Two sample t-test"
output: html_document
date: "2025-05-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

Let's compare the means of two samples using independent t-test (and paired t-test as
extra exercises)

 - Replace _ to complete expressions and return the correct answer  
 - Restart R to remove previous results using: Ctrl/Cmd + shift + F10
 - Remember - try to use Google to help you answer questions

## Import the packages

Import the `tidyverse` for `dplyr` and `ggplot2` packages

```{r}
_
```


## Two-sample t-test

`River_pH.csv` contains pH measurements between two rivers. A researcher is interested 
if the mean pH is different between the two rivers.

### Import data

Import `river_ph.csv` from the `data/` folder

```{r}
river_ph <-
  _
```

### Examine the dataset

- How many columns are there?
- How many values of River_name are there?
- How many samples in each group?

```{r}

```



### Visualise data

Visualise the data using a boxplot. Does there appear to be a difference in pH?

```{r}
river_ph %>% 
  _
```

Update the axes labels, add a title, and export the plot as a png with dimensions 100 x 100 mm to the outputs folder.

```{r}
river_ph %>% 
  _

ggsave(_)
```

### Summarise 

Obtain a data summary including
 - the number of samples
 - mean
 - standard deviation (sd)
 - standard error ( sd / sqrt(n) )
 
```{r}
river_ph_data_summary <-
  river_ph %>% 
  group_by(_) %>% 
  summarise(
    n_samples = n(),
    mean_pH = _(_),
    sd_pH = _(_),
    se_pH = sd_pH / sqrt(n_samples)
  )

river_ph_data_summary
```

### Conduct t-test

What is the null hypothesis here?

```text

```

Conduct a t.test of pH between rivers.

 - Assume equal variances for traditional t-test (var.equal = T)
 - Remember to use a formula y ~ x

```{r}
t.test(_, data = _, var.equal = T)
```

```text

```

```{r, include=F}
# If you have successfully completed the code above, try to "knit" the document"
# Use the "knit" button at the top of the script panel
```

## Additional exercises - Paired t-test

`madrid-weather-monthly-2010-2020.csv` contains daily temperature, humidity and precipitation every two years from 2010 to 2020.

A researcher wants to know if there is a difference in monthly temperature, on
average, between the year 2010 and the year 2020. We will pair the months from
each year and conduct a paired t-test.

```{r, eval=FALSE}
### Import data

madrid_weather <-
  readr::read_csv(_)

### Prepare data

# - Filter the data for the years 2010 and 2020
# - Summarise the daily data by year and month to obtain mean monthly Temperature_C

madrid_weather_monthly_2010_2020 <-
  madrid_weather %>% 
  filter(year %in% c(_)) %>% 
  group_by(_, _) %>% 
  summarise(
    Temperature_C = mean(_),
    .groups = 'drop'
  )


# Next you need to pivot the data so there are 2 columns of monthly data from 2010 and 2020


madrid_weather_monthly_2010_2020_wide <-
  madrid_weather_monthly_2010_2020 %>% 
  pivot_wider(names_from = _,
              values_from = _,
              names_prefix = 'year_')


### Plot the data

# There a different ways to do this. To save time, I've provided a two

#### Using a line graph

# - Plot monthly temperature as two coloured lines from each year
# - There should only be 2 values in the legend


madrid_weather_monthly_2010_2020 %>% 
  ggplot(aes(x = _, y = _, color = _)) +
  _

#### Boxplot of differences

# - Calculate the difference between 2020 and 2010
# - Then use a boxplot to plot the calculated difference

madrid_weather_monthly_2010_2020_wide %>% 
  mutate(temp_difference = _ - _) %>% 
  ggplot(aes(x = _)) +
  _ +
  scale_y_continuous(limits = c(-1,1),
                     breaks = NULL) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  labs(title = 'Monthly temperature difference between 2020 and 2010',
       x = 'Temperature')

# Export the plot if you wish


### Conduct a paired t-test

# What is the null hypothesis here?
_

# - Remember to provide the arguments x and y the columns of the data, and to
# - use the argument paired = TRUE
 

t.test(
  x = _,
  y = _,
  paired = _
)

# Write a short summary sentence


```



## Conclusion 

We examined

 - independent t-test
 - paired t-test
 - reviewed some data manipulation
 - reviewed some data visualisation
